Implementation of a lattice Boltzmann method for immiscible multiphase flow simulations using the level set method

### 2015-11-07 Paper/paper.tex
Paper: das meiste ist jetzt da. Results und Conclusion fehlen noch.

\subsection{Coupling of LBM and level set method}
The lattice Boltzmann method requires a boundary treatment at the interface which implements the jump conditions (\ref{jumpconditions}). We use a new boundary condition that is based on a simple bounce back type Dirichlet condition. For example, in Fig. \ref{intro_phases} for a point $x_2 \in \Omega_2$ that is next to the boundary we set
% TODO: zweites Bild einfügen mit x_1, x_2 und c_i
\begin{equation}
	f_i(x_2,t+1) = f_{i*}^+(x_2,t) + 6hf_i^*c_i \cdot \tilde{u} + R_i \text{.}
\end{equation}
Here, $\tilde{u}$ describes the linear interpolation of the velocity along the direction $c_i$, evaluated at the location $\tilde{x} = x_1 + qc_i = x_2 + (q-1)c_i$ on the interface,
$$
\tilde{u} = qu(x_2,t) + (1-q)u(x_1,t) \text{.}
$$
The additional term $R_i$ ensures the jump conditions of the normal stress and corrects the error terms that are caused by the bounce back treatment. $R_i$ is given as:
\begin{equation}
	R_i = 6 h^2 f_i^* \Lambda_i : A \text{ with } A = -q(1-q)[S] - (q - \frac12)S^{(2)} \text{.}
\end{equation}
We can rewrite the formula in the form
\begin{equation}
	R_i = 6h^2f_i^* \cdot (-q(1-q)\Lambda_i:[S] - (q-\frac12)\Lambda_i:S^{(2)} \text{,}
\end{equation}
with
\begin{equation}
	S^{(k)} = -\frac{3}{2\tau h^2} \sum_i c_i \otimes c_i(f_i-f_i^{eq})(t,x_k)
\end{equation}
\begin{equation}
	\Lambda_i:[S] = ([S]:n \otimes n)((n \cdot c_i)^2 - {|c_i|}^2/3) + 2([S]:n \otimes t_k)(n \cdot c_i)(t_k \cdot c_i) \text{.}
\end{equation}
For the calculation we use
\begin{equation}
	[S]:n \otimes n =  \frac{1}{2 \bar\mu} ([p] + 2 \sigma \kappa) - \frac{[\mu]}{\bar{\mu}} \bar{S} : n \otimes n \text{,}
\end{equation}
\begin{equation}
	[S]:n \otimes t_k = - \frac{[\mu]}{\bar{\mu}} \bar{S} : n \otimes t_k \text{.}
\end{equation}
The pressure jump is formulated by
\begin{equation}
	[p] \approx \frac{1}{3 h^2} (\rho(x_1,t)\varrho^{(1)} - \rho(x_2,t)\varrho^{(2)}) \text{.}
\end{equation}
We use now the information of the level set method for the new boundary condition. What is missing now, is a refill method. Due to the fact that the interface moves, some cells change their fluid type. To do so we use a equilibrium/non-equilibrium refill (CITE). This refill procedure consists out of four steps:
\begin{itemize}
	\item[1] Interpolate density and velocity from interior neighbors
	\item[2] Compute the corresponding equilibrium
	\item[3] Copy the non-equilibrium part from a direct interior neighbor
	\item[4] Reinitialise by adding equilibrium and non-equilibrium parts
\end{itemize}
Suppose we choose an inward pointing direction $c_i$ in a boundary point $x = x_b \in \Omega_1$, e.g. the direction of smallest angle with the surface normal. Then the density is interpolated in the following way:
\begin{equation}
	\tilde{p} = 3\rho(x-c_i,t+1) - 3\rho(x-2c_i,t+1) + \rho(x-3c_i,t+1) \text{.}
\end{equation}
To interpolate u we choose ....
The same procedure has to be used in the second fluid $\Omega_2$.

TODO: Explain the refill method. Show Fig.2 of the Th\"ommes-paper and explain the coupling of the two methods.

>>>>

\subsection{Coupling of LBM and level set method}
The lattice Boltzmann method requires a boundary treatment at the interface which implements the jump conditions (\ref{jumpconditions}). We use a new boundary condition that is based on a simple bounce back type Dirichlet condition. For example, in Fig. \ref{intro_phases} for a point $x_2 \in \Omega_2$ that is next to the boundary we set
% TODO: zweites Bild einfügen mit x_1, x_2 und c_i
\begin{equation}
	f_i(x_2,t+1) = f_{i*}^+(x_2,t) + 6hf_i^*c_i \cdot \tilde{u} + R_i \text{.}
\end{equation}
Here, $\tilde{u}$ describes the linear interpolation of the velocity along the direction $c_i$, evaluated at the location $\tilde{x} = x_1 + qc_i = x_2 + (q-1)c_i$ on the interface,
$$
\tilde{u} = qu(x_2,t) + (1-q)u(x_1,t) \text{.}
$$
The additional term $R_i$ ensures the jump conditions of the normal stress and corrects the error terms that are caused by the bounce back treatment. $R_i$ is given as:
\begin{equation}
	R_i = 6 h^2 f_i^* \Lambda_i : A \text{ with } A = -q(1-q)[S] - (q - \frac12)S^{(2)} \text{.}
\end{equation}
We can rewrite the formula in the form
\begin{equation}
	R_i = 6h^2f_i^* \cdot (-q(1-q)\Lambda_i:[S] - (q-\frac12)\Lambda_i:S^{(2)} \text{,}
\end{equation}
with
\begin{equation}
	S^{(k)} = -\frac{3}{2\tau h^2} \sum_i c_i \otimes c_i(f_i-f_i^{eq})(t,x_k) \text{,}
\end{equation}
\begin{equation}
	\Lambda_i:[S] = ([S]:n \otimes n)((n \cdot c_i)^2 - {|c_i|}^2/3) + 2([S]:n \otimes t_k)(n \cdot c_i)(t_k \cdot c_i) \text{.}
\end{equation}
For the calculation we use
\begin{equation}
	[S]:n \otimes n =  \frac{1}{2 \bar\mu} ([p] + 2 \sigma \kappa) - \frac{[\mu]}{\bar{\mu}} \bar{S} : n \otimes n \text{,}
\end{equation}
\begin{equation}
	[S]:n \otimes t_k = - \frac{[\mu]}{\bar{\mu}} \bar{S} : n \otimes t_k \text{.}
\end{equation}
Here $n$ is the normal of the interface, $t$ the tangential, $\kappa$ the curvature, and $\mu$ the dynamic viscosity.
The pressure jump is formulated by
\begin{equation}
	[p] \approx \frac{1}{3 h^2} (\rho(x_1,t)\varrho^{(1)} - \rho(x_2,t)\varrho^{(2)}) \text{.}
\end{equation}
We use now the information of the level set method for the new boundary condition. What is missing now, is a refill method. Due to the fact that the interface moves, some cells change their fluid type. We use a equilibrium/non-equilibrium refill (CITE) that consists out of four steps:
\begin{itemize}
	\item[1] Interpolate density and velocity from interior neighbors
	\item[2] Compute the corresponding equilibrium
	\item[3] Copy the non-equilibrium part from a direct interior neighbor
	\item[4] Reinitialise by adding equilibrium and non-equilibrium parts
\end{itemize}
Suppose we choose an inward pointing direction $c_i$ in a boundary point $x = x_b \in \Omega_1$, e.g. the direction of smallest angle with the surface normal. Then the density is interpolated by
\begin{equation}
	\tilde{\rho} = 3\rho(x-c_i,t+1) - 3\rho(x-2c_i,t+1) + \rho(x-3c_i,t+1) \text{.}
\end{equation}
In \cite{Thoemmes} the velocity is interpolated by using the velocity of the interface. The Ian-Mitchell toolbox does not provide a function to access this value. Therefore, we interpolate the velocity in the same way as we did for the density:
\begin{equation}
\tilde{u} = 3u(x-c_i,t+1) - 3u(x-2c_i,t+1) + u(x-3c_i,t+1) \text{.}
\end{equation}
We can now calculate the new equilibrium $f_i^{eq}(\tilde{\rho},\tilde{u})$ and add it to the non-equilibrium part that we copy from the direct neighbor, $f_i^{neq}(x-c_i,t+1)$.
The same procedure has also to be applied in the second fluid $\Omega_2$.

Finally we couple the two methods in the following way:
\begin{itemize}
	\item[1.] Create the initial interface $\Gamma_0$
	\item[2.] Run the level set code once to create the surface description for LBM
	\item[3.] Run the LBM using the data from the level set for a prescribed number of steps
	\item[4.] Pass the current velocity field to the level set code and run it for the same number of steps (see Fig. MISSING)
	\item[5.] Repeat step 3 and 4 until the end of the simulation.
\end{itemize}
